---
title: "Week_09_Homework"
author: "Jacquin Moraga"
format: html
---

## Load Libraries

Keeping it simple today
```{r}
#| warning: false
#| message: false

library(tidyverse)
library(here)
library(forcats)
```


## Load Data

Using the Intertidal Data for making the plot

```{r}
#| warning: false
#| message: false

intertidal <- read_csv(here("Week_09","Data","intertidaldata.csv"))

view(intertidal)
```


## Time to make a plot!

I wanted to make a bar graph to represent this data becuase I feel this is a goof type of plot to help visualize non-continuous data. I have seen bar graphs sued for intertidal data before which helped solidify my descision.

```{r}
#| warning: false
#| message: false

sessile <- intertidal %>%
  pivot_longer(cols = 'Mussels':'Anemone',
               names_to = "Sessile",
               values_to = "Counts") %>% #made the data in the long format so it was easier to work with
  mutate(zone = factor(Quadrat, levels = c("Low", "Mid", "High"))) #mutating the zones into factors so that they could be plotted on the x axis

sessile %>% #using this new data set I made
  drop_na() %>% #no NA's because that makes a whole new column
  ggplot(aes(x = zone, y = Counts, fill = Sessile)) + #setting the parameters for my bar graph to be the correct columns
  geom_bar(stat = "identity", position = "fill") + #we doin a bar graph cuz its my new fav, filling it by organism type
  scale_y_continuous(labels = scales::percent_format()) + #turning the y axis into percentages because I originally thought that was what the data was and then I jsut liked it so I kept it
  scale_fill_viridis_d(option = "C") + #groovy colors
  labs(title = "Sessile Organisms Coverage of Intertidal Zones",
       subtitle = "Coverage by Different Intertidal Zones",
       x = "Zone",
       y = "Percent", 
       fill = "Organism") + #giving things titles so we know what's what
  theme_bw() + #my fav theme
  theme(plot.title = element_text(size = 14),
        plot.subtitle = element_text(size = 12),
        axis.title.x = element_text(size = 10),
        axis.title.y = element_text(size = 10),
        legend.title = element_text(size = 10)) #making the text the sizes I like

ggsave(here("Week_09","Output","Homework_09_Plot.png"))

```

BEHOLD the most beautiful plot ever made.